

\chapter*{Acknowledgments}\label{ch:acknowledgments}

I would like to express my deepest gratitude to all those who have supported me throughout the process of writing this thesis.
First and foremost, I am incredibly grateful to my external supervisors, Prof. Frédéric Peschanshi and Prof. Antoine Genitrini, whose guidance, expertise, and encouragement were invaluable at every stage of this research.
Their insights into graph theory and algorithm design provided me with the foundation I needed to explore modular decomposition in depth.

A special thanks goes to my internal supervisor, Prof. Andrea Corradini, for his constant support, constructive feedback, and mentorship throughout the development of the thesis.
His attention to detail and dedication to excellence inspired me to push the boundaries of my work and ensure that the research met the highest standards.

I would also like to thank Eleni Pistiloglou, whose prior work on modular decomposition laid the groundwork for my own contributions.

I am immensely grateful to my family and friends for their unwavering support and encouragement, especially during the more challenging moments of this journey.
Their belief in me has been a constant source of motivation.
Even when they're far away from me, from another continent, they're all close to me.

A special thought to my family from another parent, my brothers and sisters from another father and another mother, as we say so well in Cameroon, to Amira, Arnold, Christelle, Maria, Mimie des Mimie (whose smile gave me lots of energy, ahahahah), Monique, Pavel, Rhynelle, Ulrich, Yannick, who throughout this work have motivated me to keep moving forward.
They have been there for me both physically and psychologically, encouraging me to believe in myself, in my abilities and in my dream.

I'd like to thank all those people who have had an impact on my life and contributed, in one way or another, to bringing me to where I am today.
A special thought to my colleagues and bosses at IperDesign who supported me and gave me the time I needed to write this report.

I am also thankful to JetBrains for offering the tools such as RustRover, which significantly aided the development and testing of my Rust implementations.

Lastly, I would like to acknowledge the Rust and C++ communities for their contributions to open-source libraries and documentation, which greatly facilitated my learning and development process.
This work would not have been possible without the collaborative spirit of the open-source community.

I dedicate my thesis to my second mother, Maman Mireille, who today is in heaven watching over us, your children.
Today you are no more, but my love for you remains.


